The analogy of the cohort-equations paradigm to the compartmental epidemic models

### 2021-05-18 note/cohort.tex
more math to the cohort work

\begin{document}
\maketitle
\linenumbers

\section{Vision}
\begin{enumerate}
\item 2 frameworks of \cite{van2002reproduction}, I call it compartment framework, and \cite{champredon2018equivalence}, I call it cohort framework, can be tied together and a mechanistic approach to go from one framework to another can be constructed.

\item 
\end{enumerate}

\section{Introduction}
Notation; we use $I'$ for the cohort-framework of $dI/d\tau$ and $\dot I$ for the compartment-framework of $dI/dt$.

A simple example of this transition is the cohort analogy of a simple SIR compartmental model where $\dot I=\beta S I/N-\gamma I$. The cohort-framework is via the cohort Eq. with the following steps
step 1: write the cohort Eq. $I'=-\gamma I$ with the initial condition $I(0)=1$, 
step 2: finding the intrinsic infectiousness kernel by integrating the cohort Eq. and solving for its time evolution, thus
$k(\tau)=\beta \exp(-\gamma \tau)$.
step 3: $R$ will be the integration of the kernel. That is $R=\beta/\gamma$. 
Note that in the comartment framework, steps 2 and 3 are combined.

In Watmough 2002's work matrix form of the I compartment $I'=(-V)I$, where I is n-by-1 and $V$ is n-by-1.

In the context of our SIR model with testing, $I^T=[I_u,I_n,I_p,I_c]$ where $T$ is the transposed operator.
And the solution is given by $I(\tau) = \exp(-V\tau) I(0)$, where $\tau$ is in the infection-time scale.
$K(\tau)=\Rnot g(\tau)= F I(\tau)$,
where $F$ is n-by-n matrix of new infections, $I(0)=[1,0,0,0]$.

$\exp(-V\tau)$ is the probability of being infected and stay infectious at time $\tau$.

$\exp(-V\tau)=\sum_{i=0}^\infty \tau^n (-V)^n/n!$.

>>>>

\begin{document}
\maketitle
\linenumbers
%%%%%%%%%%%%%%%%%%%%
\section{Vision}
\begin{enumerate}
\item 2 frameworks of \cite{van2002reproduction}, I call it compartment framework, and \cite{champredon2018equivalence}, I call it cohort framework, can be tied together and a mechanistic approach to go from one framework to another can be constructed.

\item Having the cohort framework enables one to study the strength-like and speed-like interventions. For example in our SIR model with testing and isolation, testing susciptables is a strength-like intervention and testing the infecteds' is a speed-like intervention. \todo{more context is required here.}
\end{enumerate}

%%%%%%%%%%%%%%%%%%%%
\section{Math foundation}
Notation; we use $I'$ for the cohort-framework of $dI/d\tau$, where $\tau$ is in the infection-time scale and $\dot I$ for the compartment-framework of $dI/dt$.

The matrix form of the cohort framework consists of the following 3 steps.
\begin{enumerate}[step 1]
\item
write the cohort Eq. $I'=-V I$, with the initial condition $I(0)$, where the n-by-n matrix $V$ is the flow matrix of leaving infected compartments in \cite{van2002reproduction}.

\item 
finding the intrinsic infectiousness kernel by integrating the cohort Eq. and solving for its time evolution, thus
the solution would be $I(\tau) = \exp(-V\tau) I(0)$. The kernel will be 
$K(\tau) = F I(\tau)$, where $F$ is the matrix of new infections in the compartment framework.
Note that the n-by-n matrix $\exp(-V\tau)$ is the probability of being infected and stay infectious at time $\tau$, where
$\exp(-V\tau)=\sum_{n=0}^\infty \tau^n (-V)^n/n!$.

\item
Calculating $R$ by integrating kernel $K(\tau)$.
\end{enumerate}

Note that in the comartment framework, steps 2 and 3 are combined.
%%%%%%%%%%%%%%%%%%%%
\section{examples}

{\bf Example 1, the simple SIR;}
The cohort analogy of a simple SIR compartmental model where $\dot I=\beta S I/N-\gamma I$. The cohort framework is via the cohort Eq. with the following steps
step 1: write the cohort Eq. $I'=-\gamma I$ with the initial condition $I(0)=1$,

step 2: finding the intrinsic infectiousness kernel by integrating the cohort Eq. and solving for its time evolution, thus
$k(\tau)=\beta \exp(-\gamma \tau)$.

step 3: $R$ will be the integration of the kernel. That is $R=\beta/\gamma$.

{\bf Example 2, the SIR model with testing;}
In the context of our SIR model with testing, $I^T=[I_u,I_n,I_p,I_c]$ where $T$ is the transposed operator.
And the solution is given by $I(\tau) = \exp(-V\tau) I(0)$ where $I(0)=[1,0,0,0]$.

%%%%%%%%%%%%%%%%%%%%
